---
title: 'STAT 327: Project'
author: "Dawson Kinsman and Sabastian Zuhorski"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gtools)
library(knitr)
library(kableExtra)
```

# Functions and Classes
Some functions and classes are imported from a separate file.
```{r}
source('tictactoe_functions.R')

randomSimulation = function(numSims, boardSize = 3, board = NA, df, seed = 0, colName = NA){
  set.seed(seed)
  
  winners<- replicate(numSims, randtictactoe(boardSize, board))
  table(winners)
  
  if (is.na(colName)){
    hist(winners, 
       main = paste0('Results of 1,000 ',boardSize,'x', boardSize,' Tic-Tac-Toe Games'))
    df2 <- data.frame(Winner = winners, 
                  'Board Size' = rep(paste0(boardSize,'x', boardSize), numSims))
  }
  else if (colName == 'First Move'){
    df2 <- data.frame(Winner = winners, 'Move' = rep(which(!is.na(board))), numSims)
  }
  
  df <- rbind(df, df2)
  
  return(df)
}
```
## Constant Values
```{r}
n = 1e3
```

# How does the size of the board effect the probability of Player 1 winning?
                              
First, we simulate the data using the functions we have written. We choose to run 1,000 simulations for boards with dimensions $3\times 3$, $4 \times 4$, $5 \times 5$, and $10\times 10$. 
```{r}
df <- data.frame(matrix(ncol = 2, nrow = 0))
colnames(df) <- c('Winner', 'Board Size')

df <- randomSimulation(n, 3, df = df)
df <- randomSimulation(n, 4, df = df, seed = 1)
df <- randomSimulation(n, 5, df = df, seed = 2)
df <- randomSimulation(n, 10, df = df, seed = 3)
```

We can visualize the number of times each player wins using a bar chart. 
```{r}
boards <- c('3x3','4x4','5x5','10x10')
df$Board.Size <- factor(df$Board.Size, levels = boards)
df$Winner <- factor(df$Winner, levels = c(1, -1, 0), 
                    labels = c('Player 1', 'Player 2', 'Draw'))

ggplot(df, aes(Winner)) + 
  geom_bar(aes(fill = Board.Size), position = 'dodge', color = 'black', 
           alpha = 0.75) +
  ggtitle('Winner of 1,000 Tic-Tac-Toe Games', 
          subtitle = 'Using Boards of Various Dimensions')
```

We then offer 
```{r}
results <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(results) <- c('Winner', 'Board Size', 'Probability')

for (i in boards){
  df_board <- df %>% filter(Board.Size == i)
  p1 <- length(df_board$Winner[df_board$Winner == 'Player 1']) / n; p1
  p2 <- length(df_board$Winner[df_board$Winner == 'Player 2']) / n; p2
  draw <- length(df_board$Winner[df_board$Winner == 'Draw']) / n; draw
  
  res <- data.frame(Winner = c('Player 1', 'Player 2', 'Draw'), 
                 'Board Size' = i, Probability = c(p1,p2, draw))
  results <- rbind(results, res)
}

pt <- pivot_wider(results, id_cols = Board.Size, names_from = Winner, values_from = Probability); pt
knitr::kable(pt, align = 'cccc') %>% kable_styling(full_width = F)
```

# For which first move does Player 1 have the highest probability of winning?
There are 3 general first moves, those being, 1, 4, and 5. All of the other first moves can be written as a rotation or reflection of these three
Let's simulate 1,000 games using each of the afro 

```{r}
board1 <- array(c(1, rep(NA, 8)), dim = c(3,3))
board4 <- array(c(rep(NA, 3), 1, rep(NA, 4)), dim = c(3,3))
board5 <- array(c(rep(NA,4),1,rep(NA,4)), dim = c(3,3))

df1 <- data.frame(matrix(ncol = 2, nrow = 0))
colnames(df) <- c('Winner', 'Board Size')

df1 <- randomSimulation(n, 3, board1, df = df1, seed = 0, 'First Move')
df1 <- randomSimulation(n, 3, board4, df = df1, seed = 1, 'First Move')
df1 <- randomSimulation(n, 3, board5, df = df1, seed = 2, 'First Move')
```

```{r}
df1$Move <- as.factor(df1$Move)
df1$Winner <- factor(df1$Winner, levels = c(1, -1, 0), 
                    labels = c('Player 1', 'Player 2', 'Draw'))

ggplot(df1, aes(Winner)) + 
  geom_bar(aes(fill = Move), position = 'dodge', color = 'black', 
           alpha = 0.75) +
  ggtitle('Winner of 1,000 Tic-Tac-Toe Games', 
          subtitle = 'Using Different First Moves')
```

```{r}
results <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(results) <- c('Winner', 'First Move', 'Probability')

for (i in c(1,4,5)){
  df_fmove <- df1 %>% filter(Move == i)
  p1 <- length(df_fmove$Winner[df_fmove$Winner == 'Player 1']) / n; p1
  p2 <- length(df_fmove$Winner[df_fmove$Winner == 'Player 2']) / n; p2
  draw <- length(df_fmove$Winner[df_fmove$Winner == 'Draw']) / n; draw
  
  res <- data.frame(Winner = c('Player 1', 'Player 2', 'Draw'), 
                 'First Move' = i, Probability = c(p1,p2, draw))
  results <- rbind(results, res)
}

pt <- pivot_wider(results, id_cols = First.Move, names_from = Winner, values_from = Probability); pt
knitr::kable(pt, align = 'cccc') %>% kable_styling(full_width = F)
```

# Notes
- We can check if 3 in a row by summing three entries and checking if they are 0 mod 3.
- Right now Player1 is coded as 1, Player2 is coded as -1, and a draw is 0. And I am using winner as an indicator if the game is finished. If not, then winner is NULL.

