---
title: 'STAT 327: Project'
author: "Dawson Kinsman and Sabastian Zuhorski"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gtools)
```

# Classes
Class definitions for the Monte Carlo Tree Search are shown below. 
```{r}
source('tictactoe_functions.R')
```
# Test
                              
```{r}
set.seed(0)
winners3 <- replicate(1e4, randtictactoe(3))
table(winners3)
hist(winners3, main = 'Results of 10,000 3x3 Tic-Tac-Toe Games')
```

```{r}
set.seed(1)
winners4 <- replicate(1e4, randtictactoe(4))
table(winners4)
hist(winners4, main = 'Results of 10,000 4x4 Tic-Tac-Toe Games')
```

```{r}
set.seed(2)
winners5 <- replicate(1e4, randtictactoe(5))
table(winners5)
hist(winners5, main = 'Results of 10,000 5x5 Tic-Tac-Toe Games')
```


```{r}
set.seed(3)
winners10 <- replicate(1e4, randtictactoe(10))
table(winners10)
hist(winners10, main = 'Results of 10,000 10x10 Tic-Tac-Toe Games')
```


# Notes
- We can check if 3 in a row by summing three entries and checking if they are 0 mod 3.
- Right now Player1 is coded as 1, Player2 is coded as 0, and a draw is -1. And I am using winner as an indicator if the game is finished. If not, then winner is NULL.
- For more state on a magic square: https://mathworld.wolfram.com/MagicSquare.html, we have to keep this in mind if we try to visualize any of the boards





